---
title: "BST239 HW2"
author: "Jonathan Luu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Question 8

a. Yes, it would be appropriate. 
b. If given a list of these patients, then it would be appropriate.
c. A SRS may work but other designs might be more suitable
d. If the sample is large enough, the values may be close but other designs would be better.

# Question 10

Option 3 will give the most precision as it has the largest sample size.

# Question 11

```{r}
data <- c(rep(9, 13), rep(10,35), rep(11,44), rep(12,69),
          rep(13,36), rep(14,24), rep(15,7), rep(16,3),
          rep(17,2), rep(18,5), 19, 20)
hist(data, right=F)
```

## Part 11a

No, it is not normally distributed and is right-skewed. However, since the sample is around 240 it is large enough that it should be normally distributed.

## Part 11b

```{r}
paste("The mean is", round(mean(data),3))
paste("The SE is", round(sqrt(var(data)/length(data)),3))
paste0("The 95% CI is [", 12.079-1.96*(0.124), ",", 12.079+1.96*(0.124), "]")
```

## Part 11c

The estimated sample size is 

$$
\frac{(1.96^2)(3.71)}{0.5^2} = 57
$$

# Question 12


## Part 12a

$$
\begin{aligned}
n_0 &= \frac{(1.96)^2 (0.5)^2}{0.1^2} = 96.04\\
n &= \frac{96.04}{1+96.04/580} = 82.4
\end{aligned}
$$

## Part 12b

$$
\frac{27}{120} \pm 1.96\sqrt{\frac{\frac{27}{120}(1-\frac{27}{120})}{120}} = [0.15,0.30]
$$

# Question 13

## Part 13a

$$
\begin{aligned}
&(1-\frac{745}{2700})(0.2)(0.8)\frac{1}{744}=0.00016\\
&0.2 \pm 1.96\sqrt{0.00016} = [0.175, 0.225]
\end{aligned}
$$

## Part 13b
Must be a random sample selected, else the 95% CI estimate will be biased.
